# Bananas

The farm has a combination of sugar cane and bananas. Some of the banana fields have been replanted to cane in the past few years. The fields currently in production are 1011, 101, and a newly combined 96 (old 93 + 96). Data from 2008 has been captured and will be presented in the following section.

```{r load, message=FALSE}
# Load the relevant packages
library(tidyverse)
library(lubridate)

# Read in the data using `readr` from the `tidyverse` package
production <- read_csv("data/Banana Production.csv")

# A quick look to make sure the data looks like we expect
head(production)
tail(production)

# Correct date
production$Date <- as.Date(production$Date, "%m/%d/%y")

# Add month and year columns
production$month <- floor_date(production$Date, "month")
production$year <- floor_date(production$Date, "year")

```

The fields were recorded as numbers but I would rather have the prefixed by 'f' so that I do not confuse them with production.

```{r}
# I have used the pipe function from the package dply alongside the stringr package to do this
dat<-production %>%
  mutate(Field = stringr::str_replace(Field, "93", "f93"),
         Field = stringr::str_replace(Field, "96", "f96"),
         Field = stringr::str_replace(Field, "101", "f101"),
         Field = stringr::str_replace(Field, "1011", "f1011"),
         Field = stringr::str_replace(Field, "102", "f102"))
```

Because in 2015 some banana fields were taken out and then two smaller parts were combined the nameing structure differs. To fix this I have written a few lines that account for the change







